Learning filters from markers

### 2020-03-14 texto.tex
Updates from Overleaf

@@

Let $D \subset \Z^d$ for $d > 1$ and let $\m{I}: D \to \R^b$, with $b \ge 1$, be a function that assings to each $p = (x_p, y_q) \in D$ a vector $\m{I}(p) = (I_1(p), I_2(p), \ldots, I_b(p))$.
A multi-band and multi-dimensional image is a pair ${\hat{I} = (D, \m{I})}$. We say that $D$ is the \textit{image domain}, $\m{I}(p)$ is the \textit{local feature vector} of a space element (\textit{spel}) $p \in D$, $b$ is the number of bands, and $d$ is the number of dimensions. In this work, we focus on images with $d=2$. The set of local feature vectors is called \textit{local feature space}.

@@

A multidimensioal and multi-band filter $\F$ is a pair $(\A, \mF)$ such that $\A \in D \times D$ is a adjacency relation and for all $(p, q) \in \A$, $\mF(p, q)$ assings a vector $(F_1(p, q), F_2(p, q), \ldots, F_b(p, q)) \in \R^b$ with $b \ge 1$ that satisfies the constraint that $\mF(p, q) = \mF(p, q - p)$. We can interpret $\F$ as a moving window that can be placed at any position $p \in D$. We say that $\bF = \{\F_1, \F_2, \ldots, \F_l\}$ is a \textit{filter bank} if for any $\F_i = (\A_i, \mF_i) \in \bF$ and $\F_j = (\A_j, \mF_j) \in \bF$, $\A_i = \A_j$, that is, all filters from $\bF$ has the same adjacency relation.

Now, we define the convolution operation. The convolution between an image $\I = (D, \lI)$ and a filter bank $\bF$ with $m$ filters produces a image $\J = (\pD, \lJ)$, in which, for $p \in \pD$, $\lJ(p) = (J_1(p), J_2(p), \ldots, J_m(p))$, and \[J_i(p) = \sum_{q \in \A(p)}\lI(q) \cdot {\F_i(q - p)},\] for $m = 1, 2,  \ldots, m$.

In convolution neural networks, the filters has the purpose of detecting relevant features. This happens by producing an positive value for spels which have those feaures and negative values for the ones which does not have them. This information is handle by the activation function. An activation function is a pair $\alpha$ that takes an image $\I = (D, \lI)$ as input and produce another image $\hat{A} = (D, \m{A})$ as output.  In this work, we use the \textit{rectified linear unit}, which can be defined as $A_l(p) = \max(0, I_j(p))$, as activation function.

Pooling has the purpose of aggregating nearby information and make convolutinal neural networks less sensible to small variations in data. In this work, we use only max pooling which is a operation that takes an image $\I = (D, \lI)$ with $b$ bands as input and produces an image $\hat{P} = (\pD, \m(P))$ where, for $p \in \pD$, ${\m{P}} (p) = (P_1(p), P_2(p), \ldots, P_b(p))$ such that \[P_i(p) = \max_{q \in \A(p)}{I_i(q)},\] for $i = 1, 2, \ldots, b$, and an adjacency relation $\A$.

When computing the convolution between each image of a dataset $\D$ and each filter of a filter bank, followed by ReLU activation,
at each position $p\in D$, points $\E(p)$ from the
images $\I \in \D$ may be detected at the positive (or
negative) side of a hyperplane with normal vector $\eF_l$ at the origin of $\R^{b \times |\A|}$. The distance
between $\E(p)$ and that hyperplane is proportional to $J_l(p)
=\I(p) \cdot \eF_l$, for $l \in
\{1,2,\ldots,m\}$.

@@

Batch normalization is the operation that places the origin of
$\R^{b\times |\A|}$ at the center of the cloud of points
$\E(p)$, for $p\in D$, from all images $\hat{I}\in
\D$, and then makes those clouds of points for all $p\in D$ so
hyperspherical as possible \cite{ioffe2015batch}. This can be done independently, per
position and band, by transforming each image $\I=(D_I,\lI)$
into an image $\hat{N}=(D_I,{\bf N})$, ${\bf
  N}(p)=(N_1(p),N_2(p),\ldots,N_b(p))\in \R^b$, where
\begin{eqnarray}
  N_j(p) & = & \frac{I_j(p) - \mu_j(p)}{\sigma_j(p)} \\
  \mu_j(p) & = & \frac{1}{|\D|} \sum_{\forall \hat{I}\in \D} I_{j}(p) \\
  \sigma^2_j(p) & = & \frac{1}{|\D|} \sum_{\forall \hat{I}\in \D} \left(I_{j}(p)-\mu_j(p)\right)^2
\end{eqnarray}

A convolutional neural network may consist of a sequence of
convolutional layers followed by a pattern classifier. The
convolutional layers extract features from each input image $\I$
and create a global feature vector $\gN$ at the output of the
extra batch normalization after the last convolutional layer. The
classifier may be a linear Support Vector Machine (SVM) or a
Multi-Layer Perceptron (MLP). When using a MLP, the feature space of $\gN$ has usually
its dimensions reduced by a sequence of fully connected layers before
the decision layer.

For instance, assume that ${\gN}={\gN}^{(0)}\in \R^{m\times |D|}$
and let ${\W}^{(1)}_i \in \R^{m\times |D|}$ be the synaptic weight
vector of each neuron $i=1,2,\ldots,k_1$ in a first fully connected
layer with $k_1$ neurons. The result ${\gN}^{(1)}=(N^{(1)}_1,N^{(1)}_2,\ldots,N^{(1)}_{k_1})\in \R^{k_1}$, as
$k_1 = m\times |D|$, is obtained by $k_1$ doct products with ${\W}^{(1)}_i$, $i=1,2,\ldots,k_1$, followed by activation values, such that
\begin{eqnarray}
  N^{(1)}_i & = & \max\{0,\langle {\gN}^{(0)}, {\W}^{(1)}_i\rangle\}.
\end{eqnarray}

@@

Given an image $\I = (D, \lI)$ of a dataset $\D$ such that each image has some category $\omega \in \{1, 2, \ldots, c\}$, a \textit{marker} is a pair $(p, \omega)$ where $p \in D$. We denote by $M(\I)$ a set of markers for image $\I$. For a marker $m = (p, \omega) \in M(\I)$, a \textit{patch} is the locally expanded feauture vector of $p$ with respect to some adjacency relation $\A$. We denote by $P(\I, M, \A)$ the set of patches of an image $\I$ with respect to a set of markers and an adjacency relation.

\section{Related work}

\section{The method}
Our method is based on the idea that the network designer is an expert in the application domain or knows well enough to indicate which are the relevant characteristics of the objects of interest. The designer highlights these regions by placing markers. For each marker, the extended local feature vector is taken. These vectors are clustered and the centers of the clusterizes are used as convolutional layer filters.

Let $\D$ be a dataset of images. The designer puts markers on only a subset of the images, let's say $\pDD \subseteq \D$. That is, she produces for each $\I \in \pDD$ a set of markers $\M(\I)$.  This subset of images, using the information from the markers, is used to create a set of patches \[\Pa(\pDD) = \bigcup_{\I \in \pDD}{P(\I, M, \A)},\] with respect to an adjacency relationship also defined by the designer. In this work, we use only circular adjacency relationships.

Given the set of patches $\Pa(\pDD)$ of the subset of images $\pDD$, we use a clustering algorithm to discover groups of $\Pa(\pDD)$. Each group represents a pattern (or set of patterns) highlighted by the designer through the markers. Since the center of each cluster must be a good representative of the characteristics of the group, we can use the centroid as filters to recognize these patterns in the images of $\D$.

However, for it to work properly, patches and images need to be centered at the origin. For this, we use the information from the patches to calculate the mean and standard deviation per band. If the patches are representative of images of $\pDD$ the mean and standard deviation (per band) of $\Pa$ are very close to the mean and standard deviation (per band) of $\pDD$.

Note that the convolution in a spel $p \in D$ corresponds to the doct product of the locally extended feature vecture and the vector that represents the filter. Thus, let $H$ be a hyperplane orthogonal to the vector of a $\mF$ filter. Thus, let $ H $ be a hyperplane orthogonal to the vector of a $ \mF$ filter. Vectors similar to the vector of filter $\mF$ must fall on the same side of the hyperplane as $\mF$. That is, if the vector $\E(p)$ is on the same side of $H$ as $\mF$, then the doct product between $\E (p)$ and $\mF$ is positive, and is negative otherwise. Thus, the doct product between the center of a cluster and each of the vectors of that cluster must be positive, which implies that the center is a good filter to identify the pattern described by that cluster. In the Figure (fazer a figura) we can see an example.

@@

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{0.37\textwidth}
      \includegraphics[width=\textwidth]{figures/probabilidade-fratura.png}
      \caption{Probability map for fractures.}
      \label{fig:prob-fractures}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.37\textwidth}
      \includegraphics[width=\textwidth]{figures/probabilidade-rocha.png}
      \caption{Probability map for rocks.}
    \label{fig:prob-rocks}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.37\textwidth}
      \includegraphics[width=\textwidth]{figures/probabilidade-vegetacao.png}
      \caption{Probability map for vegetation.}
  \label{fig:prob-vegetation}
  \end{subfigure}

@@

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{0.37\textwidth}
    \includegraphics[width=\textwidth]{figures/fratura.png}
    \caption{Segmentation of fractures.}
    \label{fig:segm-fractures}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.37\textwidth}
    \includegraphics[width=\textwidth]{figures/rocha.png}
    \caption{Segmentation of rocks.}
    \label{fig:segm-rocks}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.37\textwidth}
    \includegraphics[width=\textwidth]{figures/vegetacao.png}
    \caption{Segmentation of vegetation vegetation.}
  \label{fig:segm-vegetation}
\end{subfigure}

@@

We placed markers on these images in regions that we consider important to identify the coconut and non-coconut (more diverse) classes. An example of a coconut tree image with markers can be seen in the Figure.

>>>>

\section{Related Work}

@@

Let $D \subset \Z^d$ for $d > 1$ and let $\m{I}: D \to \R^b$, with $b \ge 1$, be a function that assings to each $p = (x_p, y_q) \in D$ a vector $\m{I}(p) = (I_1(p), I_2(p), \ldots, I_b(p))$.
A multi-band and multi-dimensional image is a pair ${\hat{I} = (D, \m{I})}$. We say that $D$ is the \textit{image domain}, $\m{I}(p)$ is the \textit{local feature vector} of a space element (\textit{spel}) $p \in D$, $b$ is the number of bands, and $d$ is the number of dimensions. In this work, we focus on images with $d=2$. The local feature vectors $\m{I}(p)$, of all $p\in D$, may be thought of as points of a \textit{local feature space} $\Re^b$.

@@

A multidimensioal and multi-band filter $\F$ is a pair $(\A, \mF)$ such that $\A \in D \times D$ is a adjacency relation and for all $(p, q) \in \A$, $\mF(p, q)$ assings a vector $(F_1(p, q), F_2(p, q), \ldots, F_b(p, q)) \in \R^b$ with $b \ge 1$ that satisfies the constraint that $\mF(p, q) = \mF(q - p)$. We can interpret $\F$ as a moving window that can be placed at any position $p \in D$. We say that $\bF = \{\F_1, \F_2, \ldots, \F_l\}$ is a \textit{filter bank} if for any $\F = (\A, \mF) \in \bF$ and $\F' = (\A', \mF') \in \bF$, $\A = \A'$, that is, all filters from $\bF$ has the same adjacency relation.

Now, we define the convolution operation. The convolution between an image $\I = (D, \lI)$ and a filter bank $\bF$ with $m$ filters produces a image $\J = (\pD, \lJ)$, in which, for $p \in \pD\subseteq D$, $\lJ(p) = (J_1(p), J_2(p), \ldots, J_m(p))$, and \[J_i(p) = \sum_{q \in \A(p)}\lI(q) \cdot {{\bf F}_i(q - p)},\] for $i = 1, 2,  \ldots, m$. For the sake of simplicity, we will assume that $\pD=D$.

In convolution neural networks, the filters has the purpose of detecting relevant features. This happens by producing an positive value for spels which have those features and negative values for the ones which does not have them. This information is handled by the activation function, which takes an image $\J = (D, \lJ)$ as input and produces another image $\hat{A} = (D, \m{A})$ as output.  In this work, we use the \textit{rectified linear unit}, which the elements of $\m{A}$ are defined as $A_i(p) = \max(0, J_i(p))$, $i=1,2,\ldots,m$, as activation function.

Pooling has the purpose of aggregating nearby information and makes convolutional neural networks less sensible to small variations in data. In this work, we use only max pooling which is an operation that takes an image $\J = (D, \lJ)$ with $m$ bands as input and produces an image $\hat{P} = (\pD, \m{P})$ where, for $p \in \pD\subseteq D$, ${\m{P}} (p) = (P_1(p), P_2(p), \ldots, P_m(p))$ such that \[P_i(p) = \max_{q \in \A(p)}\{I_i(q)\},\] for $i = 1, 2, \ldots, m$, and an adjacency relation $\A$. For the sake of simplicity, we will assume that $\pD=D$.

When computing the convolution between each image of a dataset $\D$ and each filter of a filter bank, followed by ReLU activation,
at each position $p\in D$, points $\E(p)$ from the
images $\I \in \D$ may be detected at the positive (or
negative) side of a hyperplane with normal vector $\mF_i$ at the origin of $\R^{b \times |\A|}$. The distance
between $\E(p)$ and that hyperplane is proportional to $J_i(p)
=\E(p) \cdot \eF_i$, for $i \in
\{1,2,\ldots,m\}$.

@@

Batch normalization is the operation that places the origin of
$\R^{b\times |\A|}$ at the center of the cloud of points
$\E(p)$, for a given $p\in D$, from all images $\hat{I}\in
\D$, and then makes those clouds of points for all $p\in D$ so
hyperspherical as possible \cite{ioffe2015batch}. This can be done independently, per
position and band, by transforming each image $\I=(D,\lI)$
into an image $\hat{N}=(D,{\bf N})$, ${\bf
  N}(p)=(N_1(p),N_2(p),\ldots,N_b(p))\in \R^b$, where, for $i\in \{1,2,\ldots,b\}$,
\begin{eqnarray}
  N_i(p) & = & \frac{I_i(p) - \mu_i(p)}{\sigma_i(p)} \\
  \mu_i(p) & = & \frac{1}{|\D|} \sum_{\forall \hat{I}\in \D} I_{i}(p) \\
  \sigma^2_i(p) & = & \frac{1}{|\D|} \sum_{\forall \hat{I}\in \D} \left(I_{i}(p)-\mu_i(p)\right)^2
\end{eqnarray}

A convolutional neural network may consist of a sequence of
convolutional layers followed by a pattern classifier. The
convolutional layers extract features from each input image $\I$
and create a global feature vector $\gN$ at the output of an
extra batch normalization after the last convolutional layer. The
classifier may be a linear Support Vector Machine (SVM) or a
Multi-Layer Perceptron (MLP). When using a MLP, the feature space of $\gN$ has usually
its dimensions reduced by a sequence of fully connected layers before
the decision layer.

For instance, assume that ${\gN}={\gN}^{(0)}\in \R^{m\times |D|}$
and let ${\W}^{(1)}_i \in \R^{m\times |D|}$ be the synaptic weight
vector of each neuron $i=1,2,\ldots,k_1$ in a first fully connected
layer with $k_1$ neurons. The result ${\gN}^{(1)}=(N^{(1)}_1,N^{(1)}_2,\ldots,N^{(1)}_{k_1})\in \R^{k_1}$, as
$k_1 = m\times |D|$, is obtained by $k_1$ inner products with ${\W}^{(1)}_i$, $i=1,2,\ldots,k_1$, followed by activation values, such that
\begin{eqnarray}
  N^{(1)}_i & = & \max\{0,\langle {\gN}^{(0)}, {\W}^{(1)}_i\rangle\}.
\end{eqnarray}

@@

Given an image $\I = (D, \lI)$ of a dataset $\D$ such that each image has some category $\omega \in \{1, 2, \ldots, c\}$, a \textit{marker} is a pair $(p, \omega)$ where $p \in D$. We denote by $\M(\I)$ a set of markers for image $\I$. For a marker $m = (p, \omega) \in \M(\I)$, a \textit{patch} is the locally expanded feature vector $\E(p)$ of $p$ with respect to some adjacency relation $\A$. We denote by $P(\I, \M, \A)$ the set of patches of an image $\I$ with respect to a set of markers and an adjacency relation.

\section{Learning filters from markers}
Our method is based on the idea that the network designer is an expert in the application domain or knows well enough to indicate which are the relevant characteristics of the objects of interest. The designer highlights these regions by placing markers. For each marker, the extended local feature vectors of its spels are taken. These vectors are clustered and the centers of those groups are used as filters for the current convolutional layer. Since each convolutional layer transforms the input feaure space of the images into a new one, the same procedure can be repeated to create the filters of the subsequents convolutional layers. This way, it is enough to specify the number of desired convolutional layers and their hyperparameters.

Let $\D$ be a dataset of images. The designer draws markers on only a subset of the images, let's say $\pDD \subseteq \D$. That is, she produces for each $\I \in \pDD$ a set of markers $\M(\I)$.  This subset of images, using the information from the markers, is used to create a set of patches \[\Pa(\pDD) = \bigcup_{\I \in \pDD}{P(\I, \M, \A)},\] with respect to an adjacency relation also defined by the designer. In this work, we use only circular adjacency relations.

Given the set of patches $\Pa(\pDD)$ of the subset of images $\pDD$, we use a clustering algorithm to discover groups in $\Pa(\pDD)$. Each group represents local patterns from one or more classes, highlighted by the designer through the markers. Since the center of each cluster must be a good representative of the characteristics of the group, we can use the group centroids as filters to enhance those patterns in the images of $\D$.

However, for it to work properly, patches and images need to be centered at the origin. For this, we use the information from the patches to calculate the mean and standard deviation per band. If the patches are representative of images in $\D$, the mean and standard deviation (per band) of $\Pa$ are very close to the mean and standard deviation (per band) in $\D$.

\begin{figure}[t]
  \begin{center}
     \includegraphics[width=0.8\linewidth]{figures/filter.png}
  \end{center}
     \caption{Filter learned from clustering.}
  \label{fig:filter}
\end{figure}

Note that the convolution in a spel $p \in D$ corresponds to the doct product of the locally extended feature vecture and the vector that represents the filter. Thus, let $H$ be a hyperplane orthogonal to the vector of a $\mF$ filter. Thus, let $ H $ be a hyperplane orthogonal to the vector of a $ \mF$ filter. Vectors similar to the vector of filter $\mF$ must fall on the same side of the hyperplane as $\mF$. That is, if the vector $\E(p)$ is on the same side of $H$ as $\mF$, then the doct product between $\E (p)$ and $\mF$ is positive, and is negative otherwise. Thus, the doct product between the center of a cluster and each of the vectors of that cluster must be positive, which implies that the center is a good filter to identify the pattern described by that cluster. In the Figure \ref{fig:filter} we can see an example.

@@

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{0.33\textwidth}
      \includegraphics[width=\textwidth]{figures/probabilidade-fratura.png}
      \caption{Probability map for fractures.}
      \label{fig:prob-fractures}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.33\textwidth}
      \includegraphics[width=\textwidth]{figures/probabilidade-rocha.png}
      \caption{Probability map for rocks.}
    \label{fig:prob-rocks}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.33\textwidth}
      \includegraphics[width=\textwidth]{figures/probabilidade-vegetacao.png}
      \caption{Probability map for vegetation.}
  \label{fig:prob-vegetation}
  \end{subfigure}

@@

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/fratura.png}
    \caption{Segmentation of fractures.}
    \label{fig:segm-fractures}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/rocha.png}
    \caption{Segmentation of rocks.}
    \label{fig:segm-rocks}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/vegetacao.png}
    \caption{Segmentation of vegetation vegetation.}
  \label{fig:segm-vegetation}
\end{subfigure}

@@

We placed markers on these images in regions that we consider important to identify the coconut and non-coconut (more diverse) classes. An example of a coconut tree image with markers can be seen in Figure \ref{fig:coco-markers} and non-coconut tree image with markers in Figure \ref{fig:non-coco-markers}.

\begin{figure}[h!]
  \begin{center}
     \includegraphics[width=0.4\linewidth]{figures/coco_markers.png}
  \end{center}
     \caption{Markers on an image which contains a coconut tree.}
  \label{fig:coco-markers}
\end{figure}

\begin{figure}[h!]
  \begin{center}
     \includegraphics[width=0.4\linewidth]{figures/non_coco-markers.png}
  \end{center}
     \caption{Markers on an image which does not contain a coconut tree.}
  \label{fig:non-coco-markers}
\end{figure}
